Naturalness in U(1) extensions of the MSSM

### 2014-12-18 paper/e6ssm-tuning-paper.tex
got rid of end of line symbols that were visiable in emacs due to mess up between editors.
Changed discussion about figure 1 for clarity.

\documentclass[preprint,amsmath,amssymb,aps,superscriptaddress,prd,showpacs,floatfix]{revtex4-1}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}
\usepackage{hyperref}

% macros
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\SuperField}[1]{\hat{#1}}
\allowdisplaybreaks
%\nofiles
% For draft, uncomment to avoid huge spaces
\raggedbottom
%
\begin{document}

@@

In these expressions $g_2$, $g'=\sqrt{3/5}g_1$, and $g_1'$ are the $SU(2)$, (non-GUT normalized) $U(1)_Y$ and $U(1)'$ gauge couplings, respectively, and $\bar{g}^2=g_2^2+g'^2$. The charges $\tilde{Q}_1$, $\tilde{Q}_2$ and $\tilde{Q}_S$ are effective $U(1)'$ charges for $H_1$, $H_2$ and $S$, respectively. Note that the $SU(2)$ dot product is defined by $A\cdot B\equiv \epsilon_{\alpha\beta}A^\alpha B^\beta=A^2B^1-A^1B^2$. The term $\Delta V_{E_6}$ contains the Coleman-Weinberg contributions to the effective potential. In the case of the $U(1)_\chi$ model, $V_{E_6}^F$ may also contain an elementary $\mu$ term contribution, as occurs in the MSSM.\\ \\
Demanding that the Higgs fields $H_1$, $H_2$ and the singlet $S$ acquire real VEVs of the form

@@

Written in the form of Eq. (\ref{eq:E6MZequation}) we see the potential new source of fine tuning alluded to above, in the form of the third term on the right-hand side. For large values of the VEV $s$, the $D$-term contributions can be quite a bit larger than $M_Z^2$. In particular, recent experimental limits \cite{Aad:2014cka} require that the $Z'$ mass be large, with bounds of $M_{Z'}\gtrsim 2.51$ TeV in $U(1)_\psi$ models and $M_{Z'}\gtrsim 2.62$ TeV in $U(1)_\chi$ models. To satisfy these limits typically requires large values of the singlet VEV $s$. For example $s\gtrsim 6$ TeV is required in the E$_6$SSM with $U(1)'=U(1)_N$, so that $|D_{H_1}|,|D_{H_2}|\gg M_Z^2$ for $E_6$ models with $\tilde{Q}_S\neq 0$. As a result the remaining terms on the right-hand side of Eq. (\ref{eq:E6MZequation}) must be tuned to cancel the resulting very large contribution to $M_Z$. Moreover, because this is a large tree level fine tuning, it may negate the increase in naturalness that is associated with having a reduced need for heavy superpartners. In $U(1)$ extended models for which $\tilde{Q}_S\neq 0$, the importance of the $Z^\prime$ mass to the fine tuning in these models can be made even clearer by writing Eq. (\ref{eq:E6MZequation}) in the form given in \cite{Athron:2013ipa},

@@

As stated above, in order to quantify the resulting fine tuning we apply the traditional Ellis-Barbieri-Giudice measure \cite{Ellis:1986yg,Barbieri:1987fn}. A specific model is characterised by a set of $n$ model parameters $\{p_i\}$ and is defined at some input scale $M_X$. For a given parameter $p$ in this set, one computes an associated sensitivity coefficient \cite{Feng:2013pwa}

@@

\section{\label{sec:results}Results}
% Our results. Plots to go in here:
%    - tuning as a function of stop mass for the MSSM and E6SSM
Many recent papers have focussed on light stops as allowing natural
SUSY and there has been a lot of theoretical effort to find SUSY
models with a $125$ GeV Higgs boson while still having light stops
and experimental efforts to search for light stops.  This is entirely
appropriate, however the fine tuning problem from the stops depends
crucially on the RG evolution from the GUT scale, since this is how
the squark masses feed into the soft Higgs masses that appear in the
EWSB conditions.

@@

% DH:: I'm not sure if this is worth including? I've put a disclaimer in the
% text, as well.
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{mssm_higgs_tuning.png}}
\caption{Scatter plot of fine tuning in the MSSM as a function of
the lightest Higgs mass, $m_{h_1}$, for the cut-off scales (from 
bottom to top) $M_X=20$ TeV, $M_X=50$ TeV, $M_X=100$ TeV and $M_X=10^{16}$ GeV.}
\label{Fig:mssmhiggstuning}
\end{center}
\end{figure}
To illustrate this, in Fig.~\ref{Fig:mssmstoptuning} we show the variation in fine
tuning when we scan over the
stop masses and mixing, with $500 \textrm{ GeV} \leq
m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and $-3810\textrm{ GeV}\leq
A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV, $50$ TeV and $100$ TeV.
The remaining parameters we fix such that at $M_{SUSY}$ they have the values $\mu=-97.5$ GeV, $B=-84.8$ GeV,
$M_1=92.1$ GeV, $M_2=95.9$ GeV, $M_3=352$ GeV, $A_b=-117.9$ GeV, $A_\tau=-7.8$ GeV, $m_{L_i}=400$ GeV, $m_{e_i}=204$ GeV,
$m_{Q_{1,2}}=438$ GeV, $m_{u_{1,2}}436$ GeV and $m_{d_i}=438$ GeV ($i=1,2,3$). By making this choice for the 
remaining parameters we ensure that fine tuning due to each is small, allowing us to see the fine tuning
impact of the stops for stop masses of less than 1 TeV. For models in which the spectrum is heavier the fine tuning
reaches a lower bound imposed by other heavier parameters as the stop masses become very small.
Note that the Higgs mass is also allowed to vary over the scan, as shown in 
Fig.~\ref{Fig:mssmhiggstuning}.
However this tuning only really becomes large for very large stop
masses.  On the other hand if the cuttoff is increased then stop masses of
several TeV can result in tuning of [insert] as shown in
Fig.~\ref{Fig:mssmstoptuning}.\\ \\

>>>>

\documentclass[preprint,amsmath,amssymb,aps,superscriptaddress,prd,showpacs,floatfix,nofootinbib]{revtex4-1}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}
\usepackage{hyperref}
\usepackage{footmisc}
% macros
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\SuperField}[1]{\hat{#1}}

 \renewcommand\footnotelayout{\fontsize{10}{9}\selectfont}

\allowdisplaybreaks
%\nofiles
% For draft, uncomment to avoid huge spaces
\raggedbottom
%
\begin{document}

@@

In these expressions $g_2$, $g'=\sqrt{3/5}g_1$, and $g_1'$ are the
$SU(2)$, (non-GUT normalized) $U(1)_Y$ and $U(1)'$ gauge couplings,
respectively, and $\bar{g}^2=g_2^2+g'^2$. The charges $\tilde{Q}_1$,
$\tilde{Q}_2$ and $\tilde{Q}_S$ are effective $U(1)'$ charges for
$H_1$, $H_2$ and $S$, respectively. Note that the $SU(2)$ dot product
is defined by $A\cdot B\equiv \epsilon_{\alpha\beta}A^\alpha
B^\beta=A^2B^1-A^1B^2$. The term $\Delta V_{E_6}$ contains the
Coleman-Weinberg contributions to the effective potential. In the case
of the $U(1)_\chi$ model, $V_{E_6}^F$ may also contain an elementary
$\mu$ term contribution, as occurs in the MSSM.\\ \\ Demanding that
the Higgs fields $H_1$, $H_2$ and the singlet $S$ acquire real VEVs of
the form

@@

Written in the form of Eq. (\ref{eq:E6MZequation}) we see the
potential new source of fine tuning alluded to above, in the form of
the third term on the right-hand side. For large values of the VEV
$s$, the $D$-term contributions can be quite a bit larger than
$M_Z^2$. In particular, recent experimental limits \cite{Aad:2014cka}
require that the $Z'$ mass be large, with bounds of $M_{Z'}\gtrsim
2.51$ TeV in $U(1)_\psi$ models and $M_{Z'}\gtrsim 2.62$ TeV in
$U(1)_\chi$ models. To satisfy these limits typically requires large
values of the singlet VEV $s$. For example $s\gtrsim 6$ TeV is
required in the E$_6$SSM with $U(1)'=U(1)_N$, so that
$|D_{H_1}|,|D_{H_2}|\gg M_Z^2$ for $E_6$ models with $\tilde{Q}_S\neq
0$. As a result the remaining terms on the right-hand side of
Eq. (\ref{eq:E6MZequation}) must be tuned to cancel the resulting very
large contribution to $M_Z$. Moreover, because this is a large tree
level fine tuning, it may negate the increase in naturalness that is
associated with having a reduced need for heavy superpartners. In
$U(1)$ extended models for which $\tilde{Q}_S\neq 0$, the importance
of the $Z^\prime$ mass to the fine tuning in these models can be made
even clearer by writing Eq. (\ref{eq:E6MZequation}) in the form given
in \cite{Athron:2013ipa},

@@

As stated above, in order to quantify the resulting fine tuning we
apply the traditional Ellis-Barbieri-Giudice measure
\cite{Ellis:1986yg,Barbieri:1987fn}. A specific model is characterised
by a set of $n$ model parameters $\{p_i\}$ and is defined at some
input scale $M_X$. For a given parameter $p$ in this set, one computes
an associated sensitivity coefficient \cite{Feng:2013pwa}

@@

\section{\label{sec:results}Results}
% Our results. Plots to go in here:
%    - tuning as a function of stop mass for the MSSM and E6SSM
As discussed in the introduction many recent papers have focussed on
light stops as allowing natural SUSY and there has been a lot of
theoretical effort to find SUSY models with a $125$ GeV Higgs boson
while still having light stops and experimental efforts to search for
light stops.  This is entirely appropriate since there are many good
reasons to expect the soft masses to be set at high energies.  However
that is not the only possibility and the fine tuning problem depends
strongly the RG evolution from the GUT scale, as the soft Higgs masses
that appear in the EWSB conditions pick up contributions from the soft
squark masses.

@@

% DH:: I'm not sure if this is worth including? I've put a disclaimer in the
% text, as well.
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{mssm_higgs_tuning.png}}
\caption{Scatter plot of fine tuning in the MSSM as a function of
the lightest Higgs mass, $m_{h_1}$, for the cut-off scales (from 
bottom to top) $M_X=20$ TeV, $M_X=50$ TeV, $M_X=100$ TeV and $M_X=10^{16}$ GeV.}
\label{Fig:mssmhiggstuning}
\end{center}
\end{figure}
To illustrate this, in Fig.~\ref{Fig:mssmstoptuning} we show the
variation in fine tuning when we scan over the stop masses and mixing,
with $500 \textrm{ GeV} \leq m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and
$-3810\textrm{ GeV}\leq A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV,
$50$ TeV and $100$ TeV.  The remaining parameters we fix such that at
$M_{SUSY}$ they have the values $\mu=-97.5$ GeV, $B=-84.8$ GeV,
$M_1=92.1$ GeV, $M_2=95.9$ GeV, $M_3=352$ GeV, $A_b=-117.9$ GeV,
$A_\tau=-7.8$ GeV, $m_{L_i}=400$ GeV, $m_{e_i}=204$ GeV,
$m_{Q_{1,2}}=438$ GeV, $m_{u_{1,2}}436$ GeV and $m_{d_i}=438$ GeV
($i=1,2,3$). By making this choice for the remaining parameters we
ensure that fine tuning due to each is small, so that we avoid washing
out the fine tuning impact of the stops when the tuning is
small \footnote{For models in which the spectrum is heavier, when the
  stop masses become very small the fine tuning reaches a lower bound
  imposed by other heavier parameters.} as can be the case when the
stop masses are less than 1 TeV.  Note that the Higgs mass is also
allowed to vary over the scan, as shown in
Fig.~\ref{Fig:mssmhiggstuning}.  However this tuning only really
becomes large for very large stop masses.  On the other hand if the
cuttoff is increased then stop masses of several TeV can result in
tuning of [insert] as shown in Fig.~\ref{Fig:mssmstoptuning}.\\ \\
